---
fig_height: 2.6
fig_width: 4
author: Joe Brew and Lucia Fernandez
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
---

---
title: Efficient assignation of spatial clusters
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# No scientific notation
options(scipen=999)


# Packages 
library(xtable)
library(knitr)
library(Hmisc)
library(raster)
library(maptools)
library(rgeos)
library(plyr)
library(dplyr)
library(sp)
library(readr)

# Basic knitr options
opts_chunk$set(comment = NA, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, cache = F)

```

```{r}
# Machine and user info
creation_date<-Sys.Date()
si <- Sys.info()
technical_message <-
  paste0('(Generated on ',
         format(creation_date, '%B %d, %Y'),
         '\nBy ',
         si['user'],
         '\nOn a ',
         si['nodename'], 
         ' ',
         si['sysname'],
         '\nMachine using R version ',
         si['release'],
         ')')
rm(si)
```

`r paste0(technical_message)`

# The problem

We have a geographical area (the district of Manhiça, for the purposes of this example), with a number of households (1,000, for the purpose of this example). We need to assign each household to a cluster of a specific size (10, for this example). Since 1 field-worker will be assigned to each cluster, we want to minimize the inter-cluster distance (ie, the distance between households in each cluster). The intra-cluster distance (ie, the distance between clusters) is irrelevant.

# The approach

We use a computational, rather than statistical, approach to solving this problem. In simple terms, our algorithm does the following.  

1. Picks a random household to begin. 
2. Picks the nearest 9 households to form a cluster.
3. Starts a new cluster, beginning with the household for which the average distance to remaining households is the lowest.
4. Picks the nearest 9 houeholds to form a cluster.
5. Repeats until all households have been assigned to a cluster.
6. Repeats steps 1-5 multiple times, so that results can be compared and the ideal configuration can be chosen.

# Walk-through

We start with the district of Maniça.

```{r, echo = TRUE}
# Get a spatialpolygonsdataframe of mozambique
moz <- getData('GADM', country = 'MOZ', level = 2)

# Plot a map of mozambique
# plot(moz)

# Subset to just include manhica
# man <- moz[moz@data$NAME_2 == 'Manhiça', ]
mag <- moz[moz@data$NAME_2 == 'Magude', ]
# Visualize
plot(mag)
```

We then read in our data (lat, lng and point id).

```{r, echo = TRUE}
## FAKE DATA
# # Create a dataframe named locations for storing our points in
# locations <- data.frame(id = 1:1000, 
#                         lat = NA,
#                         lon = NA)

## REAL DATA
# Read in bednets data
if (Sys.info()["user"] == "Lucia") {
        real_data <- '/Users/Lucia/Dropbox/MALTEM Entomology/8. Scientific Studies/3. VECE- Efficacy/4. sample_size_house_selection/House selection bednets/updated_db/households_list_reshaped_2016-04-24.csv'

        
} else {
        real_data <- '/media/joebrew/TOSHIBA/households_list_reshaped_2016-04-24.csv'

}

locations <- read_csv(real_data)

# Create columns which are named appropriately for use with our funciton
locations$lat <- locations$LAT
locations$long <- locations$LNG
locations$id <- locations$FAMILY_ID

# # Get the range for Magude
# bb <- bbox(mag)
# 
# # Populate locations
# for (i in 1:nrow(locations)){
#   
#   # Create a placeholder to indicate whether our point is in maghica
#   in_mag <- FALSE
#   
#   # If not in maghica, create a point in maghica
#   while(!in_mag){
#     # Create a point
#     lon <- sample(seq(bb[1,1],
#                       bb[1,2],
#                       length = 10000),
#                   1)
#     lat <- sample(seq(bb[2,1],
#                       bb[2,2],
#                       length = 10000),
#                   1)
#     # Put that point into a spatial dataframe
#     the_point <- data.frame(lon = lon,
#                             lat = lat)
#     coordinates(the_point) <- ~lon+lat
#     proj4string(the_point) <- proj4string(mag)
#     
#     # Check to see if it's in the boundaries of maghica
#     in_mag <- as.logical(!is.na(over(mag, the_point)))
#   }
#   # having confirmed that the point is in maghica, 
#   # pop it into the dataframe
#   locations$lon[i] <- lon
#   locations$lat[i] <- lat
# }

# Make locations a spatial points dataframe and plot
## We do this because otherwise lat and lon will not be columns any more after transforming into a geo-point data set.
locations$x <- locations$lon
locations$y <- locations$lat
locations <- data.frame(locations)
# Remove any NA points
locations<-locations[-which(is.na(locations$lat)),]
coordinates(locations) <- ~x +y
proj4string(locations) <- proj4string(mag)
plot(mag)
points(locations, 
       col = adjustcolor('red', alpha.f = 0.3), 
       pch = 3)
```

Now the heavy-lifting begins. We will do steps 1-5 (mentioned above) to test.

```{r}
# Source our function
source('cluster_optimize.R')
# run
x = cluster_optimize(times = 1,
                     cluster_size = 12,
                     sleep = 1.3,
                     plot_map = TRUE,
                     locations = locations,
                     best_only = TRUE,
                     shp=mag)

```
